---
title: "Assignment 1"
author: "Jesse Wijlhuizen (13588818), Matthijs Vogd (12425389), "
date: "26 September 2021"
output: pdf_document
fontsize: 11pt
highlight: tango
---

```{r}
library(pacman)
p_load(pwr)
p_load(tidyverse)

```
# Excercise 1.1
## a
The data seems to be normally disributed, one can project a bell curve on the
histogram and the QQplot results in a straight line. In addition the Shapiro-Wilk
test results in p-value > 0.05. Lastly the point estimate fo mu is computed:

```{r}
data=read.table(file="birthweight.txt",header=TRUE)
par(mfrow=c(1,2)); hist(data$birthweight); qqnorm(data$birthweight)
shapiro.test(data$birthweight)
mean(data$birthweight)
```

## b

```{r}
n = length(data$birthweight)
mu = mean(data$birthweight)
sd = sd(data$birthweight)
t = qt(p=0.05, df=(n -1), lower.tail=FALSE)

ci = t * sd / sqrt(n - 1); ci
lower_bound = mu - ci; lower_bound
upper_bound = mu + ci; upper_bound
```

Bounded 90% confidence interval = 2828.62  < mu < 2997.96

## c
```{r}
t.test(data, mu = 2800, alternative = "greater")
```

The 95% confidence interval doesn’t overlap with 2800, lower bound is higher. 
So expert's claim is verified.

## d
This is a one-sided test because we are only interesten in the mean being greater 
than 2800. The upper bound is irrelevant thus it can be set to infinity. This is 
why the confidence interval shifts a little bit, there is a slight chance that 
the actual mean is above the upper bound found in 1.1b.

# Excercise 1.2
## a
Point estimate for p = x/n = 140/200 = 0.7

## b
For the calculation of the standard deviation a pseudo dataset was created in R.
For the t-value 2.62 was consevatively selected from the table using 100 df.
```{r}
a = matrix(0, 60, 1)
b = matrix(1, 140,1)
d = c(a,b)
sd = sd(d) 
n = 200
mu = 0.7
t = 2.626

ci = t * sd / sqrt(n - 1); ci
lower_bound = mu - ci; lower_bound
upper_bound = mu + ci; upper_bound
```

Bounded 99% confidence interval = 0.614  < mu < 0.786

## c
For alfa = 0.1 the p-value = 0.1254 (p > 0.05), thus the null hypothesis cannot 
be rejected. For other alfa-values, the p-value remains the same, only the 
confidence interval changes. Higher alfa-values result in smaller intervals. 
This makes sense because you allow for more uncertainty when creating the interval.
```{r}

t.test(d, mu=.75, conf.level = .99)

t.test(d, mu=.75, conf.level = .999)[4]
t.test(d, mu=.75, conf.level = .95)[4]
t.test(d, mu=.75, conf.level = .90)[4]
t.test(d, mu=.75, conf.level = .80)[4]
```

# Excercise 1.3
## a)

```{r}
weather <- read.csv('../weather.txt', sep='\t')
summary(weather)
boxplot(weather$humidity) ; boxplot(weather$temperature)
```

Temperature and humidity seem to be correctly measured. No specialities have been found.

## b)

```{r}
shapiro.test(weather$humidity);shapiro.test(weather$temperature)

par(mfrow=c(1,2)); hist(weather$humidity); hist(weather$temperature) # two histograms next to each other
qqnorm(weather$humidity); qqnorm(weather$temperature) # two QQ-plots next to each other
```
As expected, temperature measurements are not normally distributed as can be seen from the QQPlot for temperature as well as the Shapiro-Wilk normality test for temperature.

## c)
```{r}


# Manually
n <- length(weather$humidity)
mu <- mean(weather$temperature)
sd <- sd(weather$temperature)
t <- qt(p=.1/2, df=59, lower.tail=FALSE)
ci <- t * sd / sqrt(n)
c(mu + ci, mu - ci)

# Using function
k <- t.test(weather$temperature, conf.level = 0.9)
k$conf.int
```

The confidence interval indicates that we can be 90\% confident that the population mean temperature belonging to this temperature sample lies between 47.49 degrees F and 57.96 degrees F.

## d)
```{r}
sd <- weather$humidity %>% sd
E <- 1
t_alpha <- qt(p=0.05/2, df=length(weather$humidity) - 1, lower.tail=FALSE)

(t_alpha * sd / E) ^ 2

```

According to the computation, 146 is the minimum sample size for a 95% confidence interval for the mean humidity such that the confidence interval has at most length 2 percentage points.

# Exercise 1.4
## a)
We are testing for homogeneity. The data points are collected using samples from different groups separately.

## b)
```{r}
austen <- read.table('../austen.txt', header=TRUE)
austen_self <- austen %>% select(-Sand2)
cs_self <- chisq.test(austen_self)
cs_self
```

As we cannot test for an $H_a$ that Austen is consistent, we cannot prove that Austen has been consistent in her texts. However, the $\chi^2$-test results do not indicate Austen was not consistent in her novels.

## c)
```{r}
austen_other <- austen %>% mutate(Austen = Sense + Emma + Sand1)
cs_other1 <- austen_other %>% select(Sand2, Austen) %>% chisq.test
cs_other1

cs_other2 <- austen_other %>% select(Sand2, Sand1) %>% chisq.test
cs_other2

cs_other3 <- austen %>% chisq.test
cs_other3
```

As we cannot test for an $H_a$ that Austen's admirer was succesfull, we will test for evidence that Austen's admirer was not successful using a $\chi^2$-test. The three different test results (admirer's part of the book vs. average of Austen texts, admirer's part of the book vs. Austen's part of the book, admirer's part of the book vs. each respective of Austen texts) do not indicate Austen's admirer was not successful.
